---
title: "Sex and respiratory system"
format: html
editor: visual
author: Beatriz Guerra
---

# Final assignment

## Aim

The aim of this final project is to study and analyze if sex has a significant effect on respiratory system structure and function. For this aim I am going to analyze the lung function in both sexes and their physical characteristics (e.g, height). I am going to use the Peru lung data base.

## Introduction

Sex-differences in the human respiratory system have previously been identified. Specifically, women have smaller lungs than men, even when matched for height, and women have smaller large conducting airways than men, even when matched for lung size. These inherent sex-differences in the structure of the respiratory system have important effects on resting pulmonary function and the integrative response to exercise. In this study, we will analyze these differences with the Peruvian data base.

## Importing data

```{r}
#import data
library(tidyverse)
library(ggplot2)
data <- read_csv("data/raw/perulung_ems.csv")
data<- data |> mutate(id=as.integer(id), 
                sex = factor(sex, levels=c(0,1), labels=c("f","m")), 
                respsymptoms=factor(respsymptoms, levels=c(0,1), labels=c("no","yes")), 
                asthma_hist=factor(asthma_hist))
data
```

## Data

```{r}
#summary
summary(data)

#head
head(data,10)

#tail
tail(data,10)

#mean height
meanheight<-data %>%
  summarise(height = mean(height, na.rm = TRUE))
meanheight

#median height by sex
medianheight<-data %>%
  group_by(sex) %>%                        
  summarise(meanheight = mean(height, na.rm = TRUE),
            Medianheight = median(height, na.rm = TRUE))
medianheight

#mean fev1
meanfev1<-data %>%
  summarise(fev1 = mean(fev1, na.rm = TRUE))
meanfev1

#median fev1 by sex
medianfev1<-data %>%
  group_by(sex) %>%                        
  summarise(meanfev1= mean(height, na.rm = TRUE),
            Medianfev1 = median(height, na.rm = TRUE))
medianheight

#mean age
meanage<-data %>%
  summarise(age = mean(fev1, na.rm = TRUE))
meanage

#median age by sex
medianage<-data %>%
  group_by(sex) %>%                        
  summarise(meanage= mean(age, na.rm = TRUE),
            medianage = median(height, na.rm = TRUE))
medianage
```

## Group proportions

```{r}
#sex vs asthma
table1 <-table(data$sex, data$asthma_hist)
table1
```

```{r}
#sex vs respiratory
table2 <-table(data$sex, data$respsymptoms)
table2

#no symptoms vs women
#no 
```

## Analysis: descriptive stadistics

```{r}


#quantile fev1
quantile(data$fev1, c(0.25,0.5,0.75))
 
 #quantile height
 quantile(data$height, c(0.25,0.5,0.75))
```

## Analysis: variables

```{r}

```

## Data visualization

```{r}

#histogram fev1
ggplot(data, aes(x=fev1)) + 
  geom_histogram()


#histogram height
ggplot(data, aes(x=height)) + 
  geom_histogram()

#histogram height, color
data %>%
  ggplot(aes(height)) +
  geom_histogram(color = "black",
                 fill = "steelblue", 
                 bins = 30) + 
  theme_bw()

#histogram fev1, color
data %>%
  ggplot(aes(fev1)) +
  geom_histogram(color = "purple",
                 fill = "green", 
                 bins = 30) + 
  theme_bw()



#box plot 


#Scatter plot

#compare height vs fev1 in both sexes
data %>%
  ggplot(aes(x = fev1, y =height, color = sex)) +
  geom_point() + 
  theme_bw()

#compere fev1 vs age in both sexes
data %>%
  ggplot(aes(x = fev1, y =age, color = sex)) +
  geom_point() + 
  theme_bw()
```

## Examples

```{r}

```
