---
title: "FEV1 and height"
format: html
author: Beatriz Guerra 
---

## Final project

This Quatro document is the final assigment of Basic Statistics and projects in R class.

## **Aim**

The aim of this assignment is to calculate is height is related to FEV1.

## Importing data: 

In this section I will import the data:

```{r}
setwd("~/project-template-main/project-template-main/data/raw")
data<-read.csv('perulung_ems.csv')
library(ggplot2)
library(tidyverse)
library(tidyverse)
library(here)
data <- read_csv("data/raw/perulung_ems.csv")
perulung_ems<-data
str(data)
data<- data |> mutate(id=as.integer(id), 
                sex = factor(sex, levels=c(0,1), labels=c("f","m")), 
                respsymptoms=factor(respsymptoms, levels=c(0,1), labels=c("no","yes")), 
                asthma_hist=factor(asthma_hist))
data<-rename(data, gender = sex)
data
```

## Analysis

In this part I will calculate the following descriptive statistics for variable 'height' using R-functions.

```{r}
summary(data)
head(data,10)
tail(data,10)


#Mean
sum(data$height, na.rm = TRUE) / 
  sum(!is.na(data$height))
124.053

sum(data$fev1, na.rm = TRUE) / 
  sum(!is.na(data$fev1))
1.594654

data %>%
  group_by(gender) %>%                        
  summarise( = mean(bill_length_mm, na.rm = TRUE)) %>%
  kable(digits = 2)




#Quartiles
quantile(data$height, c(0.25,0.5,0.75))
 25%   50%   75% 
119.9 124.0 128.0 

#Sample variance



#Sample standard deviation



#Range, interquartile range (IQR)
```

## **Data visualization**

```{r}

data
#Histogram: Height
ggplot(data, aes(x=height)) + 
              geom_histogram(color='blue', fill='purple')

#Histogram: FEV1
ggplot(data, aes(x=fev1)) + 
  geom_histogram()

#Change color
data %>%
  ggplot(aes(x=height)) +
  geom_histogram(color = "black",
                 fill = "red", 
                 bins = 30) + 
  theme_bw()

  data %>%
  ggplot(aes(x=height)) +
  geom_histogram(color = "black",
                 fill = "green", 
                 bins = 30) + 
  theme_bw()
  
  #Height-gender
data %>%
  ggplot(aes(x = gender, y =height, fill = gender)) +
  geom_boxplot() + 
  theme_bw()

#Height-FEV1
data %>%
  ggplot(aes(x = gender, y =fev1, fill = gender)) +
  geom_boxplot() + 
  theme_bw()

#Height, FEV1, gender
data %>%
  ggplot(aes(x = height, y = fev1, color = gender)) +
  geom_point() + 
  theme_bw()

```

```{r}

```

```{r}

```

```{r}

```
